Longitudinal Manifold Estimation

### 2023-02-24 paper/main.tex

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage[perpage,symbol*]{footmisc}
\usepackage{float}
\usepackage{hyperref}
\usepackage{color}  
\usepackage{tikz}

@@

@@

The notation below is used to build to a definition of principal manifolds [still need to formally write out]:
\begin{itemize}
  \item Observed data $\left\{x_i\right\}_{i=1}^I \subset \mathbb{R}^D$
  \item Associated latent parameterizations $\left\{t_i\right\}_{i=1}^I \subset \mathbb{R}^d$ [may change this notation to allow for time points]
  \item $C_{\infty}\left(\mathbb{R}^d \to \mathbb{R}^D\right) = \left\{f \in C(\mathbb{R}^d \to \mathbb{R}^D): \lim_{\|t\|_{\mathbb{R}^d} \to \infty} \|f(t)\|_{\mathbb{R}^D} = \infty\right\}$ 
  \item $\nabla f$ denotes the gradient of $f$
  \item $\nabla^{\bigotimes 2}f$ denotes the Hessian matrix of $f$
  \item $\|\nabla^{\bigotimes 2}f\|_{\mathbb{R}^{d \times d}} = \left(\sum_{i, j = 1}^{d} |\frac{\partial^2f}{\partial t_i \partial t_j}(t)|^2\right)^{\frac{1}{2}}$
  \item $\|\nabla^{\bigotimes 2}f\|_{L^2(\mathbb{R}^{d})}^2 = \sum_{l = 1}^{D}\|\nabla^{\bigotimes 2}f_l\|_{L^2(\mathbb{R}^{d})}^2 = \sum_{l=1}^{D}\int_{\mathbb{R}^d}\sum_{i,j = 1}^{d}|\frac{\partial^2f_l}{\partial t_i \partial t_j}(t)|^2dt$ represents the total squared curvature of manifold $M_f^{d}$
  \item $\mathcal{K}_{\lambda, \mathbb{P}}(f, g) = \mathbb{E}\|X - f(\pi_{g}(X))\|_{\mathbb{R}^{D}}^2 + \lambda\|\nabla^{\bigotimes 2}f\|_{L^2(\mathbb{R}^{d})}^2$, with $\mathcal{K_{\lambda, \mathbb{P}}}(f) = \mathcal{K}_{\lambda, \mathbb{P}}(f, f)$
  \item A manifold $M_{f^*}^{d}$ determined by $f^*$ is a principal manifold for random $D$-vector $X$ with compact support $\text{supp}(\mathbb{P})$ with the tuning parameter $\lambda$ if \[%
    f_{\lambda}^{*} = \argmin_{f \in \mathcal{F}(\mathbb{P})}\mathcal{K}_{\lambda, \mathbb{P}}(f)
  ,\]%
    where \[%
      \mathcal{F}(\mathbb{P}) = \left\{f \in C_{\infty} \cap \nabla^{-biot 2}L^2(\mathbb{R}^{d} \to \mathbb{R}^{D}): \sup_{x \in \text{supp}(\mathbb{P})}\|\pi_f(x)\|_{\mathbb{R}^{d}} = 1\right\}
    .\]%
  \item $\pi_{f_\lambda^{*}}(X)$ gives a $d$-dimensional parameterization of $X$
\end{itemize}

Given this definition, we now summarize the estimation method provided. This algorithm has three main steps: data reduction, fitting, and tuning. In the reduction step, the HDMDE algorithm is used to summarize the data $\left\{x_i\right\}_{i=1}^{I}$ by points $\left\{\mu_j\right\}_{j=1}^{N}$ with $N \leq I$, and each $\mu_j$ associated with weight $\theta_j$. The fitting stage of the algorithm finds estimated functions 
\[%
  \hat{f}_\lambda = \argmin_f \mathcal{K}_{\lambda, \hat{Q}_N}(f)
,\]%
where 
\[%
  \mathcal{K}_{\lambda, \hat{Q}_N}(f) = \sum_{j = 1}^{N}\theta_j \|\mu_j - f(\pi_f(\mu_j))\|_{\mathbb{R}^{D}}^2 + \lambda\|\nabla^{\bigotimes 2}f\|_{L^2(\mathbb{R}^{d})}^2
\]% 
approximates $\mathcal{K}_{\lambda, \mathbb{P}}(f)$. The tuning stage chooses $\lambda^{*} = \argmin_{\lambda > 0}(\mathcal{D}(\hat{f}_\lambda))$, where 
\[%
  \mathcal{D}(\hat{f}_\lambda) = \frac{1}{I}\sum_{i=1}^{I}\|x_i - \hat{f}_\lambda(\pi_{\hat{f}_\lambda}(x_i))\|_{\mathbb{R}^{D}}^2
.\]%

>>>>

\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage[perpage,symbol*]{footmisc}
\usepackage{float}
\usepackage{hyperref}
\usepackage{color}  
\usepackage{tikz}

@@

\doublespacing

@@

The notation below is used to build to a definition of principal manifolds [still need to formally write out]:
\begin{itemize}
  \item Observed data $\left\{x_i\right\}_{i=1}^I \subset \mathbb{R}^D$
  \item Associated latent parameterizations $\left\{t_i\right\}_{i=1}^I \subset \mathbb{R}^d$
  \item $C_{\infty}\left(\mathbb{R}^d \to \mathbb{R}^D\right) = \left\{f \in C(\mathbb{R}^d \to \mathbb{R}^D): \lim_{\|t\|_{\mathbb{R}^d} \to \infty} \|f(t)\|_{\mathbb{R}^D} = \infty\right\}$ 
  \item $\nabla f$ denotes the gradient of $f$
  \item $\nabla^{\bigotimes 2}f$ denotes the Hessian matrix of $f$
  \item $\|\nabla^{\bigotimes 2}f\|_{\mathbb{R}^{d \times d}} = \left(\sum_{i, j = 1}^{d} |\frac{\partial^2f}{\partial t_i \partial t_j}(t)|^2\right)^{\frac{1}{2}}$
  \item $\|\nabla^{\bigotimes 2}f\|_{L^2(\mathbb{R}^{d})}^2 = \sum_{l = 1}^{D}\|\nabla^{\bigotimes 2}f_l\|_{L^2(\mathbb{R}^{d})}^2 = \sum_{l=1}^{D}\int_{\mathbb{R}^d}\sum_{i,j = 1}^{d}|\frac{\partial^2f_l}{\partial t_i \partial t_j}(t)|^2dt$ represents the total squared curvature of manifold $M_f^{d}$
  \item $\mathcal{K}_{\lambda, \mathbb{P}}(f, g) = \mathbb{E}\|X - f(\pi_{g}(X))\|_{\mathbb{R}^{D}}^2 + \lambda\|\nabla^{\bigotimes 2}f\|_{L^2(\mathbb{R}^{d})}^2$, with $\mathcal{K_{\lambda, \mathbb{P}}}(f) = \mathcal{K}_{\lambda, \mathbb{P}}(f, f)$
  \item A manifold $M_{f^*}^{d}$ determined by $f^*$ is a principal manifold for random $D$-vector $X$ with compact support $\text{supp}(\mathbb{P})$ with the tuning parameter $\lambda$ if \[%
    f_{\lambda}^{*} = \arg \min_{f \in \mathcal{F}(\mathbb{P})}\mathcal{K}_{\lambda, \mathbb{P}}(f)
  ,\]%
    where \[%
      \mathcal{F}(\mathbb{P}) = \left\{f \in C_{\infty} \cap \nabla^{-biot 2}L^2(\mathbb{R}^{d} \to \mathbb{R}^{D}): \sup_{x \in \text{supp}(\mathbb{P})}\|\pi_f(x)\|_{\mathbb{R}^{d}} = 1\right\}
    .\]%
  \item $\pi_{f_\lambda^{*}}(X)$ gives a $d$-dimensional parameterization of $X$
\end{itemize}

Given this definition, we now summarize the estimation method provided. This algorithm has three main steps: data reduction, fitting, and tuning. In the reduction step, the HDMDE algorithm is used to summarize the data $\left\{x_i\right\}_{i=1}^{I}$ by points $\left\{\mu_j\right\}_{j=1}^{N}$ with $N \leq I$, and each $\mu_j$ associated with weight $\theta_j$. The fitting stage of the algorithm finds estimated functions 
\[%
  \hat{f}_\lambda = \arg\min_f \mathcal{K}_{\lambda, \hat{Q}_N}(f)
,\]%
where 
\[%
  \mathcal{K}_{\lambda, \hat{Q}_N}(f) = \sum_{j = 1}^{N}\theta_j \|\mu_j - f(\pi_f(\mu_j))\|_{\mathbb{R}^{D}}^2 + \lambda\|\nabla^{\bigotimes 2}f\|_{L^2(\mathbb{R}^{d})}^2
\]% 
approximates $\mathcal{K}_{\lambda, \mathbb{P}}(f)$. The tuning stage chooses $\lambda^{*} = \arg\min_{\lambda > 0}(\mathcal{D}(\hat{f}_\lambda))$, where 
\[%
  \mathcal{D}(\hat{f}_\lambda) = \frac{1}{I}\sum_{i=1}^{I}\|x_i - \hat{f}_\lambda(\pi_{\hat{f}_\lambda}(x_i))\|_{\mathbb{R}^{D}}^2
.\]%
